---
title: "PROJECT ON games DATASET"
output:
  html_document:
    df_print: paged
---

# Business Understanding
Do data analysis on games data set  so as to understand the relationship between the columns and predict the future values.

# Data
This data was gotten from this site:




# Visual Exploratory Descriptive Analysis

## Import packages
```{r}
library (tidymodels)
library(visdat)
library(forecast)
```

Those are packages that are in the tidymodels which we are going to use in this project.

## Import dataset
```{r}
data <- games
glimpse(data)
```

Our data set has 1,512 and 14 columns




## Explore the data class structure visually
```{r}
vis_dat(data)
```
Most of the coulmns in our data set are character data type.

## convert all remaining character variables to factors 
```{r}
# convert all remaining character variables to factors 
data <- data %>% 
           mutate(across(where(is.character), as.factor))
vis_dat(data)
```
There is no character data type  column.

## Visualizing missing data
```{r}
vis_miss(data, sort_miss = TRUE)
```

## Viewing missing dataset 
```{r}
is.na(data) %>% colSums()
```
The only coulmns with missing data are Team,Rating and Summary.

## Fill missing data with mode

```{r}
# Replace missing values with mode
data_filled <- data %>%
  mutate_all(~ ifelse(is.na(.), Mode(., na.rm = TRUE), .))

# Mode function to calculate mode
Mode <- function(x, na.rm = FALSE) {
  if (na.rm) {
    x <- x[!is.na(x)]
  }
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Verify if missing values have been filled with the mode
is.na(data) %>% colSums()

```
Now there is no missing data.




## 1.Bubble chart of game ratings and number of reviews
```{r}
ggplot(data, aes(x = Rating, y = as.numeric(gsub("K", "", `Number of Reviews`)), size = Rating)) +
  geom_point() +
  xlab("Rating") +
  ylab("Number of Reviews") + theme_classic()
```

Most games had a review of 4 and relatively they were around 500.




## Box plot of game ratings
```{r}
ggplot(data, aes(y = Rating)) +
  geom_boxplot() +
  ylab("Rating")
```
There are many outliers in the boxplot of game ratings.

## Scatter plot of number of reviews vs. rating


## Histogram of game ratings
```{r}
ggplot(data, aes(x = Rating)) +
  geom_histogram(binwidth = 0.1, fill = "red", color = "white") +
  xlab("Rating") +
  ylab("Frequency")
```
The game ratings has normal distribution.

## Bar chart of game ratings
```{r}
ggplot(data, aes(x = Title, y = Rating)) +
  geom_bar(stat = "identity") +
  xlab("Game Title") +
  ylab("Rating")
```
## Non-Visual Exploratory Data Analysis

## Top rows of our data set
```{r}
head(data)
```

## Bottom rows of our data set
```{r}
tail(data)
```

From the two tables we can say our data is not biased.
## Structure of our data set
```{r}
glimpse(data)
```

Our data set has just numeric and factor data type columns as data cleaning.
## Statistical summary of our data set
```{r}
skimr::skim(data)
```

## Create a new data frame that contains only the games that have been played at least 10,000 times
```{r}
data %>%
  filter("Total Plays" >= 10000)
```
## Create a new data frame that contains only the games with a rating of 4 or higher
```{r}
games_4 <- data %>%
  filter(Rating >= 4)

games_4
```
These  are games with rating of 4 and above.


# Data wrangling
Here are some data wrangling done in the above:
* Filling missing data with mode.
* Convert all character columns to factor.


## Convert columns with numeric values to numeric type
```{r}
data <- data %>%
  mutate(across(c(Rating, `Times Listed`, `Number of Reviews`, Plays, Playing, Backlogs, Wishlist), as.numeric))
```


## Convert 'Release Date' column to Date type
```{r}
data$`Release Date` <- as.Date(data$`Release Date`, format = "%b %d, %Y")
```

## Extract year and month from 'Release Date'
```{r}
# Extract year from 'Release Date'
data$Year <- lubridate::year(data$`Release Date`)

#Extract month from 'Release Date'
data$Month <- lubridate::month(data$`Release Date`, label = TRUE)
```

## Split 'Genres' column into separate genre columns
```{r}
data <- data %>%
  separate_rows(Genres, sep = "', '", convert = TRUE) %>%
  pivot_wider(names_from = Genres, values_from = Genres, values_fn = length, values_fill = 0, names_prefix = "Genre_")
```

## Remove unnecessary columns
```{r}
data <- data %>%
  select(-c(`...1`, Summary, Reviews))
```

## Rename columns to remove spaces and special characters
```{r}
# Rename columns to remove spaces and special characters
colnames(data_filled) <- make.names(colnames(data_filled), unique = TRUE)
```

## Convert K values to numeric in columns with "K" suffix
```{r}
#  Convert K values to numeric in columns with "K" suffix
data_filled[, c("Times.Listed", "Number.of.Reviews", "Plays", "Playing", "Backlogs", "Wishlist")] <- lapply(data_filled[, c("Times.Listed", "Number.of.Reviews", "Plays", "Playing", "Backlogs", "Wishlist")], function(x) as.numeric(gsub("K", "", x)))
```

## Create a new column called Total Plays that is the sum of the Plays and Playing columns
```{r}
data_filled$"Total Plays" <- data$Plays + data$Playing
head(data_filled$"Total Plays")
```



## Create a new column called Average Rating that is the average of the Rating column.
```{r}
data_filled$"Average_Rating" <- mean(data$Rating)
```

## Group the data by Genre and calculate the total number of games in each genre
```{r}
data_filled %>%
  group_by(Genres) %>%
  summarise(Count = n())
```



## See the data after data wrangling
```{r}
head(data_filled)
```

```{r}
glimpse(data_filled)
```

That is our new data frame after data wrangling.

# Visual analysis(Time series analysis using ARIMA model)


## Convert the data to a time series
```{r}
ts_data <- ts(data_filled$Rating)
```

## Fit an ARIMA model
```{r}
arima_model <- auto.arima(ts_data)
```

## Forecast future values
```{r}
forecast_data <- forecast(arima_model, h = 12)
```


## Visualize the time series and forecast
```{r}
ggplot() +
  geom_line(aes(x = time(ts_data), y = ts_data, color = "Actual")) +
  geom_line(aes(x = time(forecast_data$mean), y = forecast_data$mean, color = "Forecast")) +
  xlab("Time") +
  ylab("Rating") +
  ggtitle("Time Series Analysis using ARIMA") +
  theme_minimal()
```

The plot shows the actual values and predicted values. These data show random variation; There are no patterns or cycles.
